The Galactic Midplane Is Not a Plane: Implications for Dynamical Analysis with {\em Gaia} Data and Beyond

### 2019-05-04 paper/ms.tex
Remove volume density argument, move def of midplane to section intro

We now consider a second source of systematic offsets in the Galactocentric
coordinate system: intrinsic variations of the midplane across the disk. We
use two sets of simulations of Milky Way-mass galaxies to estimate the size of
this effect. We will also briefly discuss a rough estimate of the midplane
variation derived from the observed velocity distribution of the Milky Way,
but defer a detailed treatment to future work.

@@

The softening lengths used in the simulations can affect the ability to
resolve the very thinnest planar structures, which in turn can affect how much
the density-based midplane varies as a function of azimuth. The Milky Way's
dense, star-forming gas disk is thought to have a scale height of about
$40\,\pc$ ($3\text{-}4$ times the cold gas softening length;
\citealt{2019ApJ...871..145A}) and the thin stellar disk a scale height of
about 300 pc (30 times the stellar softening length;
\citealt{2008ApJ...673..864J}). We therefore expect that resolution effects
may still be affecting the scale heights of these components in the
simulations, especially the cold gas. Indeed, the stellar scale heights of the
simulated galaxies are equal to or larger than the Milky Way's while the gas scale
heights are significantly larger (although the proper basis comparison is less
clear in the case of the gas; the quoted value for the Milky Way comes from
studies of high-mass star-forming regions). As noted above, the midplanes
defined by gas and stars can be tilted with respect to one another as well,
precluding extending the precision of the gas midplane definition to the
stellar component. In this work we specifically consider azimuthal variations
in the midplane \emph{as defined by the stellar mass density}, which will
subsequently be lower in the simulations than in the Milky Way: within 200 pc of the
Sun the measured gas and stellar volume densities in the Milky Way are roughly
comparable at $\sim 0.04\ \Msun\
\unit{pc}{-3}$, while in the simulations they are a factor of a few lower (see
Table 3 of \citealt{2018arXiv180610564S}). We thus expect that the degree of
midplane variation we find in the simulated systems is likely to be
\emph{less} than in the real Milky Way, since on the smallest scales the simulations
are \emph{smoother} than the real Galaxy.

>>>>

We now consider a second source of systematic offsets in the Galactocentric
coordinate system: intrinsic variations of the midplane across the disk. In
this work we specifically consider azimuthal variations in the midplane as
defined by the stellar mass density. We use two sets of simulations of Milky
Way-mass galaxies to estimate the size of this effect.

@@

The softening lengths used in the simulations can affect the ability to
resolve the very thinnest planar structures, which in turn can affect how much
the density-based midplane varies as a function of azimuth. The Milky Way's
dense, star-forming gas disk is thought to have a scale height of about
$40\,\pc$ ($3\text{-}4$ times the cold gas softening length;
\citealt{2019ApJ...871..145A}) and the thin stellar disk a scale height of
about 300 pc (30 times the stellar softening length;
\citealt{2008ApJ...673..864J}). We therefore expect that resolution effects
may still be affecting the scale heights of these components in the
simulations, especially the cold gas. Indeed, the stellar scale heights of the
simulated galaxies are equal to or larger than the Milky Way's while the gas scale
heights are significantly larger (although the proper basis comparison is less
clear in the case of the gas; the quoted value for the Milky Way comes from
studies of high-mass star-forming regions). As noted above, the midplanes
defined by gas and stars can be tilted with respect to one another as well,
precluding extending the precision of the gas midplane definition to the
stellar component.
